\section{Introduction}
Large language models (LLMs) excel at reasoning, coding, and multimodal tasks. However, due to the high hardware requirements (RAM, GPU, etc.) they demand, these models are usually run only on cloud-based infrastructures or very advanced local systems. This creates a significant bottleneck for system developers seeking to reduce cloud dependency, ensure data privacy, and avoid high operational costs.

As a result of these limitations, small language models (SLMs) are emerging as a strong alternative. Their ability to run on local machines provides a significant advantage in terms of data security, and their operational costs are much lower compared to LLMs. The ability to fine-tune them for specific domains, their easy integration with explicit control modules, and their collaborative operation structure make them quite attractive. However, SLMs cannot provide the broad expertise offered by a single massive model (like GPT-4o). They still fall short in tasks requiring logical reasoning, deep contextual understanding, and extensive world knowledge. Therefore, reducing the issue to a simple "SLM or LLM?" competition is no longer a fair or comprehensive approach. Instead, current research revolves around questions such as in which areas a single SLM is sufficient, when an LLM is needed, when a hybrid or agent-based architecture should be used, and how a customized SLM system can outperform a single LLM.

The existing literature provides concrete examples of how these architectural choices are shaped by the dynamics of the field. For example, in the telecommunications sector, the Symbiotic Agents study has managed to significantly reduce error rates and resource consumption by using hybrid and multi-agent systems \cite{chatzistefanidis2025survey}. In medicine, the Self-Reflective Chest X-Ray Report Generation study combined three different AI models in a single pipeline---a self-reflective agent (LLaMA-8B) and an LLM judge---to create a hybrid structure that maximized diagnostic accuracy and minimized error rates \cite{han2025chest}. Similarly, in cybersecurity, Pietrzak and colleagues solved a complex fake report detection problem using only a fine-tuned single SLM (TinyLlama) and detection models (TinyBERT, DistilBERT) \cite{pietrzak2025evaluation}. They achieved this success with just standard consumer-grade GPU hardware, achieving higher detection rates compared to large models like GPT-4 \cite{pietrzak2025evaluation}.

These developments provide practical alternatives for scenarios where relying on a single large language model is insufficient or excessively costly. Moreover, research in this area not only focuses on basic accuracy rates but also aims to provide tangible improvements in computational load, operational cost, and latency. Researchers have clearly demonstrated which types of hybrid designs work for specific tasks and what kinds of architectures are effective in real-world conditions.

This review investigates how hybrid models converge across different fields, when LLMs should play a more active role, and in which cases control can be delegated to SLMs. It examines in which specific domains fine-tuned SLMs perform best and where they naturally fall short. By analyzing recurring hybrid methods in the literature, the article reveals scenarios where SLMs can realistically replace or complement large language models. Since the scope is fundamentally built on the resource-performance trade-off, this review places special emphasis on hybrid and multi-agent systems.

The main contributions of this review are as follows:
\begin{itemize}
    \item It provides a synthesized overview of 54 studies covering cloud-based monolithic LLMs, hybrid SLM-LLM architectures, the performance of fine-tuned SLMs versus large models, and multi-agent workflows.
    \item It frames the literature around four distinct practical research questions: Performance, Efficiency, Architecture, and Domain Suitability.
    \item It creates an analytical review matrix that reveals the commonalities and differences among the studies, mapping which specific research question each study addresses.
\end{itemize}