\section{Summary of the Studies}
Before detailing the specific findings related to the research questions, this section provides a structural overview of the 54 primary studies included in this review. During the data extraction phase, 16 key characteristics were systematically captured for each study to facilitate comparative analysis. Each extraction field is defined below:

\begin{itemize}
    \item \textbf{Study ID:} A unique identifier (S1--S54) assigned to each study for streamlined referencing throughout the text and appendix.
    \item \textbf{Year:} Indicates the year the study was formally published or released as a preprint.
    \item \textbf{SLM(s):} Lists the specific small language model(s) evaluated, fine-tuned, or deployed within the proposed system.
    \item \textbf{LLM(s):} Specifies the large language model(s) utilized either as a performance baseline, a teacher for distillation, or a cloud-based fallback.
    \item \textbf{Architecture (Arch.):} Describes the overarching structural design of the system, such as hybrid edge-cloud, multi-agent ensemble, or retrieval-augmented generation.
    \item \textbf{Orchestration Method:} Specifies the exact mechanism used to coordinate models and tasks, such as token-level routing, prompt-guided fusion, or majority voting.
    \item \textbf{Agent Count:} Indicates the number of distinct agents, roles, or expert models operating within the proposed system.
    \item \textbf{Task Type:} Details the specific downstream application, domain, or computational problem evaluated in the study.
    \item \textbf{Dataset:} Identifies the specific benchmarks, open-source datasets, or proprietary clinical/industrial data used to train and evaluate the system.
    \item \textbf{Metrics:} Outlines the primary evaluation criteria used to assess model performance, efficiency, and resource consumption.
    \item \textbf{LLM Baseline:} Reports the baseline performance score of the standalone large language model.
    \item \textbf{SLM Baseline:} Reports the baseline performance score of the standalone small language model.
    \item \textbf{Accuracy:} Presents the final performance or accuracy metric achieved by the proposed hybrid or SLM-centered architecture.
    \item \textbf{Latency (ms):} Captures the inference speed, time-to-first-token, or overall time delay reported by the authors.
    \item \textbf{Cost (\$):} Highlights any reported financial savings regarding API usage or computational reductions relative to full LLM inference.
    \item \textbf{Energy:} Details power consumption, Joules-per-token, or carbon footprint implications, if measured by the authors.
    \item \textbf{Key Findings:} Summarizes the core takeaway and primary contribution of the study regarding SLM-LLM orchestration.
    \item \textbf{Limitations:} Highlights the constraints, failure modes, data dependencies, or hardware bottlenecks identified by the authors.
\end{itemize}

Because the full comparative analysis table is located in the online appendix, the mapping of Study IDs to their respective bibliographic citations is provided below for quick reference:

\textbf{Primary Studies Mapping:} 
S1: \cite{tigani2025cpu}, S2: \cite{hao2024hybrid}, S3: \cite{xie2025fusing}, S4: \cite{alabbasi2024teleoracle}, S5: \cite{cao2025enersafe}, S6: \cite{lee2024orchestrallm}, S7: \cite{han2025commentagent}, S8: \cite{aydin2025integrated}, S9: \cite{demir2025dpatransllm}, S10: \cite{ji2025raprag}, S11: \cite{pacheco2025flight}, S12: \cite{dorfner2024comparing}, S13: \cite{zhu2024ofslms}, S14: \cite{jeong2023testtime}, S15: \cite{relora2025}, S16: \cite{chatzistefanidis2025survey}, S17: \cite{han2025chest}, S18: \cite{pietrzak2025evaluation}, S19: \cite{diagnosis2025}, S20: \cite{resiliomate2025}, S21: \cite{arabic2025}, S22: \cite{prompt2024}, S23: \cite{feedback2025}, S24: \cite{tcm2025}, S25: \cite{psa2026}, S26: \cite{vte2024}, S27: \cite{multiagent2025}, S28: \cite{compos2024}, S29: \cite{networkrl2025}, S30: \cite{alabbasi2024teleoracle}, S31: \cite{academicrag2024}, S32: \cite{rethinkllm2025}, S33: \cite{fedhlm2025}, S34: \cite{energytoken2025}, S35: \cite{arasum2025}, S36: \cite{hlmea2025}, S37: \cite{debateagents2025}, S38: \cite{phishing2024}, S39: \cite{transport2025}, S40: \cite{textnex2025}, S41: \cite{olfactory2025}, S42: \cite{staticanalysis2025}, S43: \cite{zhu2024distilling}, S44: \cite{colbertrag2024}, S45: \cite{oh2025uncertainty}, S46: \cite{citer2025}, S47: \cite{sentensemble2026}, S48: \cite{qwqrouting2025}, S49: \cite{egi2025}, S50: \cite{itemdiff2025}, S51: \cite{bioagents2025}, S52: \cite{amarasinghe2025fipo}, S53: \cite{shao2025dot}, S54: \cite{mixeddistill2024}.

The comprehensive data extraction spreadsheet can be accessed at: \url{https://bit.ly/SLM-LLM-Data-2026}.

\section{Findings}

\subsection{Corpus-Level Trends Before the RQ Analysis}
Before answering the four research questions directly, three corpus-level observations are worth emphasizing.

First, the field is now primarily a systems literature rather than a pure model-scaling literature. Although a few studies still compare standalone small and large models directly, the dominant trend is to embed them inside a larger control strategy. Hybrid edge-cloud pipelines, retrieval-centered systems, expert routers, task decomposers, self-reflective workflows, and fusion modules are now common design choices. In other words, recent work increasingly treats model scale as one variable in a broader systems problem.

Second, hybridization is motivated by different constraints in different domains. In telecom and networking, the main concerns are deployment, context fidelity, and latency budgets. In healthcare and scientific tasks, the dominant concerns are reliability, interpretability, and domain specificity. In industrial monitoring, multimodal hybridization is driven by the need to combine coarse global understanding with fine-grained local detail. In optimisation and education, hybrid methods often serve as a way to recover reasoning quality from small models through feedback, routing, or targeted prompting.

Third, the evidence base is uneven. Accuracy is almost always reported; compute, latency, and cost are not. This matters because a large share of the literature frames hybridization as an efficiency solution, yet only a minority of studies expose sufficiently comparable numbers to test those claims rigorously. The result is a literature that often proves that hybridization \emph{can} work, but less often proves how much it costs in practice under standardized conditions.
\subsection{RQ1: Comparative Performance of Hybrid and Multi-Agent Systems}

The reviewed evidence shows that both hybrid and multi-agent architectures can outperform standalone small-model baselines, but their advantages emerge through different forms of coordination rather than parameter scaling alone. Hybrid systems achieve their strongest gains when computation is selectively distributed between local small models and stronger remote models. Token-level edge--cloud collaboration improves small-model performance without requiring full-query escalation: one study reports LLM-comparable quality on GSM8K at only 25.8\% of full-cloud cost by using TinyLlama 1.1B or Llama-2-7B as edge-side drafting models and Llama-2-70B as a cloud verifier (S2). Similarly, routing only about 7\% of low-confidence tokens from a 0.5B on-device model to a cloud-hosted Qwen2.5-32B-Instruct model yields over 60\% performance improvement over the local-only setup while preserving more than 80\% cost reduction relative to full LLM inference (S46). This pattern is reinforced by uncertainty-aware hybrid inference, where opportunistic skipping reduces uplink transmissions and LLM computations by 45.93\%, reaches up to 97.54\% of LLM inference accuracy, and delivers 2.54$\times$ faster token throughput than hybrid inference without skipping (S45). At the sub-task level, structured task decomposition and cloud--edge allocation reduce average reasoning time by 66.12\% and API cost by 83.57\% while maintaining reasoning accuracy comparable to the best baselines (S53). In domain-specific settings, hybrid orchestration can also outperform larger general-purpose models: a fine-tuned Phi-2 (2.7B) combined with retrieval, reranking, hybrid search, and long-context support achieves 84.20\% accuracy, exceeding GPT-4o mini (82.05\%) and GPT-4o (80.80\%) on telecom question answering (S4).

Multi-agent systems, by contrast, show their comparative strength when performance depends on role specialization, critique, verification, and distributed reasoning. A three-agent Phi-3-based architecture performs on par with human experts in conceptual genomics tasks, although its gains weaken on more complex code-generation workflows, indicating that multi-agent benefits remain task-sensitive (S51). In reasoning-intensive question answering, a Proponent--Opponent--Arbitrator structure built on compact models such as Qwen2.5 0.5B and DeepSeek-R1 1.5B surpasses both single-agent reasoning and vanilla LLM baselines with only 1.8$\times$ inference-time overhead (S37). Likewise, assistant--checker collaboration supported by memory and feedback loops enables small-model multi-agent systems to approach or exceed larger baselines across several tasks: LeptoConnect-7B outperforms GPT-3.5-turbo by 163.1\% on database tasks, while LeptoConnect-1.8B reaches 81.6\% of GPT-4's score on knowledge graph tasks, and an enhanced TinyLlama-1.8B improves ROUGE-L by 63.4\% over CodeLlama-7B in code correction (S20). Overall, the findings indicate that hybrid systems are most effective when the main challenge is balancing accuracy with latency and cost through adaptive routing, whereas multi-agent systems are most effective when the task benefits from explicit decomposition of reasoning roles and iterative verification. Thus, the comparative performance advantage of both architecture families is driven primarily by orchestration strategy, not by model size alone.
\subsection{RQ2: Efficiency, Cost, Latency, and Energy Trade-offs}

The Appendix A data reveals that while efficiency motivates the adoption of hybrid and multi-agent systems, the actual impact on cost, latency, and energy depends heavily on the orchestration method. Hybridization is not a universal efficiency solution; it introduces a trade-off space where computational bottlenecks are often shifted rather than eliminated.

Regarding inference cost and token consumption, architectures employing selective escalation or task decomposition drastically reduce reliance on massive standalone LLMs. Study S46 achieves an over 80\% cost reduction compared to full LLM inference by routing fewer than 7\% of generated tokens to the cloud. Similarly, S2 uses a token-level draft-verification method to match the quality of a 70B parameter LLM while consuming only 25.8\% to 31.2\% of standard cloud inference costs. Query-level decomposition yields similar savings: S53 cuts average API costs by 83.57\% via parallel reasoning. Computational savings are also heavily emphasized through token reduction. Study S33 achieves a 95\% reduction in LLM token transmissions via federated uncertainty-aware routing, and S27 eliminates redundant multi-round communications through an SLM-generated directed acyclic graph, reducing token consumption by 35\% to 70\%. In dialogue state tracking, S6 highlights that a retrieval-based router reduces overall computational FLOPs by approximately 62\%.

Conversely, the impact on latency is highly polarized. In optimal scenarios, relying on lightweight models accelerates throughput. A fine-tuned SLM in a hybrid classification pipeline (S32) operates at 7.41 ms on a GPU, making it over 20 times faster than its larger counterpart. Furthermore, S45 shows that uncertainty-aware skipping in an edge-cloud setup provides 2.54$\times$ faster token throughput. However, orchestration mechanisms frequently introduce new latency bottlenecks. Planners and retrievers add non-trivial temporal overhead; the adaptive retrieval planner in S10 takes 1.42 seconds compared to 0.95 seconds for naive retrieval, and S4 explicitly identifies cross-encoder reranking as its primary latency bottleneck. Multi-agent collaboration also inherently multiplies inference cycles. Study S37 reports a 1.8$\times$ inference-time overhead for adversarial debate, and complex agentic workflows in S22 can push average response latencies to 107 seconds. Processing time is further impacted by hardware and modality constraints, such as constrained local hardware (S11), dual-path video sampling (S5), CPU-only zero-shot ensembles taking upwards of 859 seconds (S47), and multi-turn key-value cache recomputation at the edge (S48).

Energy consumption remains the least documented metric in the reviewed corpus, though the few studies that measure it provide critical systems insights. Study S16 demonstrates the resource efficiency of small models by showing that symbiotic SLM agents reduce the VRAM footprint by 99.9\% while saving 44.43\% of physical resource blocks in radio access networks. Memory architecture also aids efficiency; study S20 shows that a dual-scale memory system reduces computational redundancy by 26.4\%, inherently lowering the energy required for repeated edge computations. However, study S34 provides a groundbreaking caveat: forcing a small model to execute complex reasoning workflows, such as Chain-of-Thought, drastically inflates its energy consumption per query. In specific contexts, querying a massive standalone LLM without Chain-of-Thought is actually more energy-efficient than forcing an SLM through a lengthy reasoning prompt. This finding underscores that smaller parameter counts do not automatically guarantee lower energy footprints if the prompting strategy artificially extends the generation length.

\subsection{RQ3: Orchestration and Routing Mechanisms}
The corpus contains a broad set of orchestration mechanisms, but they can be organized into a small number of recurring families. Across the included studies, performance differences are explained less by absolute model size than by how effectively the system retrieves evidence, routes uncertainty, and divides labor across components.

\paragraph{Token-level collaboration.}
Token-level orchestration is most visible in verification-oriented hybrid systems. Study S2 uses the SLM as a drafter and the LLM as a verifier on difficult tokens. Study S45 extends this pattern with uncertainty-aware skipping so that likely acceptable SLM tokens can bypass the remote model. Study S46 reaches a related design point by routing only low-confidence tokens for external inference. Taken together, these studies show that token-level collaboration provides fine-grained control over answer quality, but it can also introduce communication and synchronization overhead when escalation rates increase.

\paragraph{Query-level or example-level routing.}
Query-level routing appears when the system selects a processing path for the full input rather than for individual tokens. Study S6 exemplifies retrieval-based expert routing at the dialogue level, while Study S32 shows that instance-level model selection can also be effective in customer-review analysis. These designs are most useful when ``easy'' and ``hard'' cases can be separated in a semantically meaningful way before full generation begins.

\paragraph{Retrieval-centered orchestration.}
RAG is a major control mechanism in the corpus even when the architecture is classified as hybrid rather than purely RAG. Studies S4, S10, S31, and S44 use retrieval not only to expand context but also to constrain the action space of a smaller model. In these studies, retrieval acts as a coordination layer that narrows ambiguity, grounds the answer in domain evidence, and reduces reliance on raw model scale alone.

\paragraph{Planner-based decomposition and agentic workflows.}
Task decomposition is increasingly important in the recent corpus. Study S53 uses a task decomposer, dependency graph, and plug-and-play adapter to allocate subtasks between local and remote models. Study S52 uses a feedback-integrated workflow that gathers structured feedback from multiple perspectives before revising the prompt of the smaller model. Related role-based decomposition also appears in Study S11, Study S20, and Study S51. Across these studies, the common idea is that small models remain viable when the workflow can be partitioned into specialist stages rather than handled as a single undifferentiated generation step.

\paragraph{Fusion and multimodal coordination.}
Not all orchestration is text-only. Study S5 combines coarse-grained and fine-grained visual semantics through prompt-guided fusion, while Study S3 fuses language-model embeddings with graph-based features for molecular-property prediction. In these settings, orchestration is best understood as structured fusion across heterogeneous sources of evidence.

These patterns support a broader interpretation of orchestration. It is not merely an implementation detail added after model selection. Across the included studies, orchestration is often the main source of improvement. The strongest systems define when the SLM should act alone, when evidence should be retrieved, and when a larger model or another specialist should be involved. This helps explain why the same small model may perform weakly in monolithic form yet remain competitive when embedded in a more deliberate control architecture.

\subsection{RQ4: Domain Suitability and Application-Specific Constraints}

The extracted data reveals that the choice between hybrid and multi-agent systems is rarely domain-agnostic. The overarching constraints of the operational environment---whether they be strict privacy mandates, rigid latency budgets, or the need for iterative logical verification---dictate the orchestration method. Across the 54 primary studies, distinct architectural preferences emerge across four core applied domains.

\paragraph{Healthcare, Clinical, and Scientific Reasoning}
In fields with strict privacy constraints (e.g., HIPAA compliance) or high-stakes accuracy requirements, hybridization is driven by the need to keep data local while grounding the SLM in verifiable facts. Consequently, local SLMs fine-tuned on clinical corpora and augmented by Retrieval-Augmented Generation (RAG) are the dominant architecture (Studies S17, S21). For example, Study S17 presents a hybrid framework for medical report generation that successfully improved patient readability across 6th to 18th-grade levels without losing clinical rigor by maintaining local detection. When evaluating radiology reports, Study S12 demonstrated that few-shot prompting allowed open-source models to statistically match the performance of massive models like GPT-4, achieving a Micro F1 score of 0.98 versus 0.97. This reliance on local ensembles and structured tabular transformations extends heavily into diagnostic classifications, ranging from herbal formulas to cardiovascular outcome probabilities (Studies S19, S24). Distillation also proves highly effective in this domain; Study S35 showed that distilling knowledge from a massive 30B parameter model into a lightweight SLM yields a highly accurate, privacy-preserving tool for multilingual medical summarization. In complex scientific domains, such as molecular property prediction, Study S3 demonstrated that a 1.5B parameter DeepSeek model fused with Graph Convolutional Networks (GCN) outperformed larger 7B and 8B variants, proving that local models can master highly specialized scientific vernacular when structurally augmented.

\paragraph{Software Engineering and Educational Assessment}
Tasks requiring rigid syntax, iterative refinement, and logical debugging benefit significantly from multi-agent or feedback-driven SLMs rather than zero-shot massive LLMs. In pedagogical text generation, Study S1 showed that a local 1.7B SLM orchestrated via a Markov Decision Process achieved a quality score of 4.17, which is directly comparable to the 4.15--4.31 range achieved by cloud-based LLMs. For code correction and database operations, Study S20 utilized specialized small models integrated with dual-scale memory systems to reduce computational redundancy by 26.4\% and semantic mistakes by 34.8\%. Automated assessment also benefits from structural orchestration, where highly structured report components allow smaller models to out-compete larger baselines (Study S23). Furthermore, orchestration frameworks frequently assign specific roles to SLMs to mimic the developer-reviewer dynamic and generate high-fidelity code offline (Studies S13, S51, S52). Study S51 successfully utilized a three-agent system to generate bioinformatics code that approaches human-expert baselines, while Study S52 deployed a Feedback-Integrated Prompt Optimiser (FIPO) to iteratively revise SLM outputs based on explicit execution feedback. Finally, for code static analysis, Study S42 found that while massive LLMs demonstrate a clear advantage for broad type inference, smaller specialized models often suffice when integrated properly into the analysis pipeline.

\paragraph{Telecom and Networking}
Telecommunications architectures prioritize strict latency budgets, bandwidth allocation, and domain-specific vernacular, making hybrid edge-cloud systems the standard. The TeleOracle system evaluated in Study S4 demonstrated that an SLM generator could achieve an 81.20\% accuracy in telecom domain QA---a 30\% improvement over the base Phi-2 model---effectively matching GPT-4 while keeping data within telecom network constraints. This is reinforced by Study S44, which found that combining LoRA fine-tuning with a highly accurate ColBERT retrieval mechanism allows small models to effectively parse and answer highly technical telecom jargon. Beyond text generation, latency optimization and bandwidth reduction are critical drivers for hybridization (Studies S29, S45, S53). Study S29 integrated lightweight sequence predictions with reinforcement learning to enable proactive bandwidth scheduling, significantly outperforming traditional reactive methods. Similarly, Studies S45 and S53 utilized uncertainty-aware speculative inference and sub-task decomposition to dynamically bypass cloud LLMs, drastically reducing uplink transmissions and cutting average API costs by up to 83.57\%.

\paragraph{Edge Deployment, Industrial, and Cyber-Physical Systems}
Real-world interactions require physical hardware constraints, such as CPU-only limitations or embedded edge devices, to govern model orchestration. In hybrid edge-cloud setups for complex reasoning, Study S2 emphasizes that only approximately 3\% of tokens are deemed ``hard'' enough to require cloud LLM verification, drastically saving edge compute resources. In mission-critical physical environments, distributed multi-agent architectures are relied upon to process raw operational data for geospatial and aerospace decision support (Studies S11, S39). Study S11 highlights that locally embedded SLMs (e.g., Mistral 7B) can achieve accuracy comparable to cloud-based GPT-4o for specialized cockpit operations. In industrial video monitoring, Study S5 utilizes a dual-path hybrid framework to decouple visual information: the local SLM processes coarse-grained, uniform frames for rapid assessment, and only triggers the larger multimodal LLM to process keyframes when a complex, localized safety anomaly is detected. This modality-spanning orchestration extends into emerging cyber-physical tasks, such as cross-modal olfactory understanding in Study S41, where fine-tuned, heavily quantized multimodal SLMs process local sensor data to bridge the gap between physical inputs and language without remote computation.

\subsection{Cross-RQ Synthesis}
Viewed together, the four research questions suggest that the literature is converging toward a specific systems principle: hybrid language-model design works best when the large model is treated as a selectively invoked capability rather than the default engine for every token, query, or subtask.

The first implication is architectural. Strong results are associated with explicit control layers such as routers, planners, retrievers, or fusion modules. This means that future progress is unlikely to come only from making SLMs slightly larger or LLMs slightly cheaper. It will also depend on better confidence estimation, task decomposition, retrieval selection, and cross-model scheduling.

The second implication is methodological. Because accuracy dominates reporting, many studies overstate practical efficiency. The corpus already contains cases where stronger orchestration clearly reduces cost or latency, but it also contains cases where coordination shifts the bottleneck rather than removing it. A mature evaluation protocol should therefore report resource metrics as first-class outcomes rather than optional extras.

The third implication is application-oriented. The most successful SLM-centered deployments are not generic substitutes for LLMs. They are domain-adapted systems in which the small model is grounded, specialized, or protected by escalation logic. This is why telecom, structured reasoning, radiology labeling, optimisation formulation, and industrial monitoring are so prominent in the positive findings of the corpus.
