\section{Threats to Validity}
This review has several threats to validity that should be stated explicitly.

\textbf{Internal validity.} Although the review follows a PRISMA-guided workflow, study identification, screening, and extraction still involve researcher judgment, especially for borderline papers in which the SLM component or empirical contribution was not equally central. The representative PDF inspection used to deepen the narrative adds a second judgment layer because not every included study was read in the same depth. To reduce this risk, borderline screening decisions were discussed through team consensus, and the synthesis remained tied to the predefined research questions and the same extraction matrix throughout.

\textbf{Construct validity.} Core constructs such as ``SLM'', ``hybrid architecture'', ``multi-agent system'', and ``efficiency'' are not standardized across the literature. A single hybrid label may refer to token-level verification, retrieval-grounded generation, planner-based decomposition, or multimodal fusion, while efficiency may mean latency, FLOPs, API cost, bandwidth reduction, or energy use. This limits direct comparability because studies grouped under the same high-level category may differ substantially in mechanism. To partially limit this, the review separates broad architecture labels from extracted orchestration mechanisms and discusses recurring mechanism families explicitly.

\textbf{Conclusion validity.} Cross-study inference is limited by the uneven and inconsistent reporting of evaluation outcomes. Accuracy is reported in most papers, but latency, cost, and energy appear only in a minority, often with different units, hardware assumptions, or deployment settings, which prevents formal meta-analysis and strong quantitative claims about efficiency gains across the corpus. Publication and reporting bias may further favor positive hybrid results, so the extraction also recorded reported limitations, and the synthesis emphasizes uneven reporting and task-sensitive trade-offs rather than universal gains.

\textbf{External validity.} The corpus is based on a Web of Science-centered search and a bounded review window, which improves transparency but may miss relevant studies from other databases, preprint platforms, or fast-moving venues. Although WoS covers many major peer-reviewed journals and conference proceedings, the field evolves quickly and the current corpus is heavily concentrated in 2025 and in a limited set of application domains. Therefore, the conclusions should be generalized cautiously beyond the covered time period, tasks, and deployment settings.
