\section{Conclusion}
This paper presented a systematic literature review of 54 studies on collaborative SLM--LLM systems across hybrid, multi-agent, ensemble, and retrieval-centered architectures, treating monolithic LLMs as comparison baselines rather than as a primary study class. The review was structured around four practical questions: comparative performance, efficiency, orchestration, and domain suitability. Across these four axes, the main pattern is consistent: SLMs are most competitive when they are supported by retrieval, routing, uncertainty estimation, task decomposition, prompt optimisation, or multimodal fusion rather than used as isolated replacements for larger models.

The findings also show that the literature is stronger in architectural innovation than in evaluation discipline. Many studies report accuracy improvements, but far fewer report latency, cost, or energy in a form that supports fair comparison. As a result, the current evidence is sufficient to show that hybridization can work, but often insufficient to show when it is the better deployment choice under realistic resource constraints. The review also suggests that the strongest positive results are concentrated in narrower or more structured settings such as telecom question answering, selected healthcare tasks, optimisation, and industrial monitoring, while open-ended reasoning and broad general-purpose use remain less convincing for SLM-centered systems.

The limited energy-related evidence in the reviewed literature supports a narrower conclusion than broad efficiency claims often imply. Direct energy measurement is rare, and the clearest energy-centered result shows that when small models are pushed into long Chain-of-Thought traces, their per-query energy use can exceed that of a larger model prompted more directly~\cite{energytoken2025}. Other studies in the corpus report only proxy improvements, such as a 26.4\% reduction in computational redundancy through dual-scale memory~\cite{resiliomate2025}, which may indicate lower resource demand but not verified energy savings. Future work should therefore treat energy as a first-class evaluation outcome and report energy-per-token, reasoning length, hardware configuration, and escalation behavior under comparable conditions. In that sense, the open question is not whether SLM-centered systems can be efficient in principle, but when their orchestration actually translates into measured energy savings in practice.
